Evolution, functional diversity, regime shifts, hysteresis, and nonlinearity; ecosystem responses to environmental change.

### 2021-07-18 experiments/experiment old/experiment 1/experiment 1.Rmd
new experiment 1 sorted

```{r}
ssfind_minimum_abundances <- rep(0, 3)
names(ssfind_minimum_abundances) <- c("CB", "PB", "SB")
ssfind_simulation_duration <- default_sim_duration
ssfind_simulation_sampling_interval <- ssfind_simulation_duration
ssfind_event_interval <- ssfind_simulation_duration
grid_num_a <- 1000 ## number of a_0 values
a_Os <- 10^seq(-7, -1, length=grid_num_a) ## sequence of a_0 values
grid_num_N <- 2 ## number of N values
initial_CBs <- 10^seq(0, 10, length=grid_num_N) ## sequence of N values
initial_PBs <- 1e8 ## not varied
initial_SBs <- 1e8 ## not varied
# next line creates all possible combinations
ss_expt <- expand.grid(N_CB = initial_CBs,
                      N_PB = initial_PBs,
                      N_SB = initial_SBs,
                      a_O = a_Os)

@@

```{r}
var_expt <- run_ss_var_experiment()
saveRDS(var_expt, here("experiments/experiment 1/data/ss_data_latest.RDS"))

@@

@@

var_expt$pars[[1]]
dd <- var_expt$ss_res[[1]]
dd1 <- filter(dd, PB_1<(-0.0001))
dd1$a_O

a_Os <- a_Os[355]
ss_expt <- expand.grid(N_CB = initial_CBs,
                      N_PB = initial_PBs,
                      N_SB = initial_SBs,
                      a_O = a_Os)

var_expt_master <- var_expt
var_expt <- var_expt[1,]
var_expt_test <- run_ss_var_experiment()
res <- var_expt_test$ss_res[[1]]
#saveRDS(var_expt, here("experiments/experiment 1/data/ss_data.RDS"))

sim_number <- 1
log10_a <- log10(a_Os[355]) ## very slowly goes anoxic
#log10_a <- log10(a_Os[354]) ## very slowly goes anoxic
#log10_a <- log10(a_Os[356]) ## very very very  slowly goes anoxic
#log10_a <- log10(a_Os[357]) ## does not go anoxic

default_log10a_series <- c(log10_a, log10_a)
initial_state <- new_initial_state(num_CB_strains,
                                   num_PB_strains,
                                   num_SB_strains,
                                   values = "bush_ssfig3")
initial_state[grep("CB_", names(initial_state))] <- 10^10/num_CB_strains
sim_res_novar <- run_simulation(parameter_values = var_expt$pars[[sim_number]],
                          initial_state = initial_state)
plot_dynamics(sim_res_novar)
#ggsave(here("simulations/expt2/figures/switching_novar.pdf"), width = 10)

>>>>

```{r}
default_sim_duration <- 1000000
ssfind_minimum_abundances <- rep(0, 3)
names(ssfind_minimum_abundances) <- c("CB", "PB", "SB")
ssfind_simulation_duration <- default_sim_duration
ssfind_simulation_sampling_interval <- ssfind_simulation_duration
ssfind_event_interval <- ssfind_simulation_duration
grid_num_a <- 1000 ## number of a_0 values
a_Os <- 10^seq(-7, -1, length=grid_num_a) ## sequence of a_0 values
grid_num_N <- 2 ## number of N values
initial_CBs <- 10^seq(0, 10, length=grid_num_N) ## sequence of N values
initial_PBs <- 1e8 ## not varied
initial_SBs <- 1e8 ## not varied
# next line creates all possible combinations
ss_expt <- expand.grid(N_CB = initial_CBs,
                      N_PB = initial_PBs,
                      N_SB = initial_SBs,
                      a_O = a_Os)

@@

```{r}
#var_expt <- run_ss_var_experiment()
#saveRDS(var_expt, here("experiments/experiment 1/data/ss_data_latest.RDS"))

@@

ss_exp_result <- var_expt
result_index <- 1

@@

I (Owen) found that the sampling interval had an effect on the stability of the simulation. If the sampling interval was long, then in some rare cases (see below) the odesolver failed, with negative abundances occuring. I think this is due to abundances becoming very small, and then the computer having trouble with precision. I guess that when a sample is taken, the abundance is somehow altered if it is very low, probably by some rounding.

```{r}
var_expt$pars[[1]]
dd <- var_expt$ss_res[[1]]
dd1 <- filter(dd, PB_1<(-0.0001))
dd1$a_O

ss_expt_master <- ss_expt
ss_expt <- ss_expt_master[abs(ss_expt_master$a_O - 1.336984e-05)<1e-10,]

var_expt_master <- var_expt
#var_expt <- var_expt[1,]
var_expt_test <- run_ss_var_experiment()
res <- var_expt_test$ss_res[[1]]

test1 <- ss_by_a_N(ss_expt, var_expt$pars[[1]])
x <- ss_expt[2,]
param <- var_expt$pars[[1]]
get_final_states_a_N(x, param)
ssfind_parameters <- param
ssfind_simulation_sampling_interval <- 1000
## now run inside the function "get_final_states_a_N"
simres1 <- simres
ssfind_simulation_sampling_interval <- 5000
## now run inside the function "get_final_states_a_N"
simres2 <- simres  # this fails

## now run inside the function "get_final_states_a_N"
plot_dynamics(simres2)

ggplot() +
  geom_line(data = simres1$result,
             mapping = aes(x = time, y = log10(PB_1))) +
  geom_point(data = simres2$result,
             mapping = aes(x = time, y = log10(PB_1))) +
  xlim(c(0, 250000))

ccc <- simres2$result

simres2$result$PB_1
simres2$result$time

log10_a <- log10(ss_expt$a_O[1]) ## very slowly goes anoxic
#log10_a <- log10(a_Os[354]) ## very slowly goes anoxic
#log10_a <- log10(a_Os[356]) ## very very very  slowly goes anoxic
#log10_a <- log10(a_Os[357]) ## does not go anoxic

default_dynamic_model <- bushplus_dynamic_model
default_event_definition <- event_definition_1
default_event_interval <- ssfind_simulation_duration
default_noise_sigma <- 0
default_minimum_abundances <- ssfind_minimum_abundances
default_sim_duration <- ssfind_simulation_duration
default_sim_sample_interval <- ssfind_simulation_duration
#initial_pars_from <- "bush_ssfig3"

default_log10a_series <- c(log10_a, log10_a)
initial_state <- new_initial_state(num_CB_strains,
                                   num_PB_strains,
                                   num_SB_strains,
                                   values = "bush_ssfig3")
initial_state[grep("CB_", names(initial_state))] <- 10^10/num_CB_strains
sim_res_novar <- run_simulation(parameter_values = var_expt$pars[[1]],
                          initial_state = initial_state)
plot_dynamics(sim_res_novar)

simulation_result <- sim_res_novar
every_n <- 1

chk <- sim_res_novar$result

sim_res_novar$result %>%
  ggplot() +
  geom_line(mapping = aes(x = time, y = PB_1))

#ggsave(here("simulations/expt2/figures/switching_novar.pdf"), width = 10)

### 2021-07-18 experiments/experiment old/experiment 1/negative abundance investigation.Rmd
new experiment 1 sorted

---
title: "Experiment 1"
author: "Owen Petchey"
date: "6/25/2021"
output:
  pdf_document: default
  html_document:
    keep_md: yes
---

@@

@@

```{r}
default_sim_duration <- 80000
```

```{r}
sim_number <- 1
default_log10a_series <- c(-1, -7, -7)
initial_state <- new_initial_state(num_CB_strains,
                                   num_PB_strains,
                                   num_SB_strains,
                                   values = "bush_ssfig3")
initial_state[grep("CB_", names(initial_state))] <- 10^10/num_CB_strains
sim_res_novar <- run_simulation(parameter_values = var_expt$pars[[sim_number]],
                          initial_state = initial_state)
plot_dynamics(sim_res_novar)
#ggsave(here("simulations/expt2/figures/switching_novar.pdf"), width = 10)
```

@@

```{r}
options(mc.cores = 12)
```

```{r}
var_expt <- run_ss_var_experiment()
saveRDS(var_expt, here("experiments/experiment 1/data/ss_data_latest.RDS"))

```{r}
var_expt <- readRDS(here("experiments/experiment 1/data/ss_data_latest.RDS"))

```{r}
result_index2 <- num_div_treatment_levels
p2  <- plot_ss_result1(var_expt,
                result_index = result_index2,
                filename_prefix = NULL,
                save_image_file = FALSE)
p2

```{r}
p_overlay <- plot_ss_result2(var_expt[result_index1,]$ss_res[[1]],
                             var_expt[result_index2,]$ss_res[[1]],
                             xlims = c(-7, -1))
p_overlay

## negative abundance investigation

var_expt$pars[[1]]
dd <- var_expt$ss_res[[1]]
dd1 <- filter(dd, PB_1<(-0.0001))
dd1$a_O

a_Os <- a_Os[355]
ss_expt <- expand.grid(N_CB = initial_CBs,
                      N_PB = initial_PBs,
                      N_SB = initial_SBs,
                      a_O = a_Os)

sim_number <- 1
log10_a <- log10(a_Os[355]) ## very slowly goes anoxic
#log10_a <- log10(a_Os[354]) ## very slowly goes anoxic
#log10_a <- log10(a_Os[356]) ## very very very  slowly goes anoxic
#log10_a <- log10(a_Os[357]) ## does not go anoxic

default_log10a_series <- c(log10_a, log10_a)
initial_state <- new_initial_state(num_CB_strains,
                                   num_PB_strains,
                                   num_SB_strains,
                                   values = "bush_ssfig3")
initial_state[grep("CB_", names(initial_state))] <- 10^10/num_CB_strains
sim_res_novar <- run_simulation(parameter_values = var_expt$pars[[sim_number]],
                          initial_state = initial_state)
plot_dynamics(sim_res_novar)
#ggsave(here("simulations/expt2/figures/switching_novar.pdf"), width = 10)

>>>>

---
title: "negative abundance investigation"
author: "Owen Petchey"
date: "7/16/2021"
output: html_document
---

@@

```{r}
options(mc.cores = 4)
```

```{r}
a_Os[355]

```{r eval = FALSE}
var_expt <- readRDS(here("experiments/experiment 1/data/ss_data.RDS"))

```{r eval = FALSE}
result_index2 <- num_div_treatment_levels
p2  <- plot_ss_result1(var_expt,
                result_index = result_index2,
                filename_prefix = NULL,
                save_image_file = FALSE)
p2

```{r eval = FALSE}
p_overlay <- plot_ss_result2(var_expt[result_index1,]$ss_res[[1]],
                             var_expt[result_index2,]$ss_res[[1]],
                             xlims = c(-7, -1))
p_overlay

@@

```{r}
var_expt <- readRDS(here("experiments/experiment 1/data/ss_data.RDS"))
```

find the negative abundances

```{r}
var_expt$pars[[1]]
dd <- var_expt$ss_res[[1]]
dd1 <- filter(dd, PB_1<(-0.0001))
dd1$a_O
```

```{r}
sim_number <- 1
log10_a <- log10(a_Os[355]) ## very slowly goes anoxic
#log10_a <- log10(a_Os[354]) ## very slowly goes anoxic
#log10_a <- log10(a_Os[356]) ## very very very  slowly goes anoxic
#log10_a <- log10(a_Os[357]) ## does not go anoxic

default_log10a_series <- c(log10_a, log10_a)
initial_state <- new_initial_state(num_CB_strains,
                                   num_PB_strains,
                                   num_SB_strains,
                                   values = "bush_ssfig3")
initial_state[grep("CB_", names(initial_state))] <- 10^10/num_CB_strains
sim_res_novar <- run_simulation(parameter_values = var_expt$pars[[sim_number]],
                          initial_state = initial_state)
plot_dynamics(sim_res_novar)
#ggsave(here("simulations/expt2/figures/switching_novar.pdf"), width = 10)
```

```{r eval = FALSE}
a_Os <- 10^seq(-7, -1, length=grid_num_a) ## sequence of a_0 values
a_Os <- a_Os[355]
ss_expt <- expand.grid(N_CB = initial_CBs,
                      N_PB = initial_PBs,
                      N_SB = initial_SBs,
                      a_O = a_Os)

@@

### 2021-07-18 manuscript/article.Rmd
new experiment 1 sorted

```{r ss_novar, out.width="1\\linewidth", include=TRUE, fig.align="center", fig.cap=c("Steady states of the ecosystem with no trait diversity, i.e. identical trait values for the strains within a functional group."), echo=FALSE}
CBSBPB_data <- readRDS(here("experiments/experiment 1/data/ss_data.RDS"))
result_index1 <- 1
p1  <- plot_ss_result1(CBSBPB_data,
                result_index = result_index1,
                filename_prefix = NULL,
                save_image_file = FALSE)
p1

@@

@@

```{r ss_var, out.width="1\\linewidth", include=TRUE, fig.align="center", fig.cap=c("Steady states of the ecosystem with diversity in all functional groups"), echo=FALSE}
result_index2 <- 20
p2  <- plot_ss_result1(CBSBPB_data,
                result_index = result_index2,
                filename_prefix = NULL,
                save_image_file = FALSE)
p2

@@

```{r ss_var2, out.width="1\\linewidth", include=TRUE, fig.align="center", fig.cap=c("Steady states of the ecosystem without any diversity (solid line) and with diversity in all three functional groups."), echo=FALSE}
p_overlay <- plot_ss_result2(CBSBPB_data[result_index1,]$ss_res[[1]],
                             CBSBPB_data[result_index2,]$ss_res[[1]],
                             xlims = c(-7, -1))
p_overlay
```

@@

```{r CB_var1, out.width="1\\linewidth", include=TRUE, fig.align="center", fig.cap=c("Steady states of the ecosystem with diversity in only the cyanobacteria functional group."), echo=FALSE}
CB_data <- readRDS(here("experiments/experiment 2/data/ss_data.RDS"))
p3  <- plot_ss_result1(CB_data,
                result_index = result_index2,
                filename_prefix = NULL,
                save_image_file = FALSE)
p3
```

```{r CB_var2, out.width="1\\linewidth", include=TRUE, fig.align="center", fig.cap=c("Steady states of the ecosystem without any diversity (solid line) and with diversity in only the cyanobacteria functional group."), echo=FALSE}
p_overlay <- plot_ss_result2(CB_data[result_index1,]$ss_res[[1]],
                             CB_data[result_index2,]$ss_res[[1]],
                             xlims = c(-7, -1))
p_overlay
```

@@

```{r SBPB_var1, out.width="1\\linewidth", include=TRUE, fig.align="center", fig.cap=c("Steady states of the ecosystem with diversity in only the two sulphur bacteria functional groups."), echo=FALSE}
SBPB_data <- readRDS(here("experiments/experiment 3/data/ss_data.RDS"))
p6  <- plot_ss_result1(SBPB_data,
                result_index = result_index2,
                filename_prefix = NULL,
                save_image_file = FALSE)
p6

@@

```{r SBPB_var2, out.width="1\\linewidth", include=TRUE, fig.align="center", fig.cap=c("Steady states of the ecosystem without any diversity (solid line) and with diversity in only the sulphur bacteria functional group."), echo=FALSE}
p_overlay <- plot_ss_result2(CB_data[result_index1,]$ss_res[[1]],
                             SBPB_data[result_index2,]$ss_res[[1]],
                             xlims = c(-7, -1))

p_overlay
```

>>>>

```{r}
## find various combintions of diversity
var_expt <- readRDS(here("experiments/experiment 4/data/ss_data.RDS"))

var_expt_levels <- var_expt[,1:6]

no_diversity <- which(rowSums(abs(var_expt_levels))==0)
max_diversty_all <- which(max(rowSums(abs(var_expt_levels))) ==
                            rowSums(abs(var_expt_levels)))
max_only_CB_diversity <- which(max(rowSums(abs(var_expt_levels[,1:2]))) ==
                            rowSums(abs(var_expt_levels[,1:2])) &
                              rowSums(abs(var_expt_levels[,3:6]))==0)
max_only_SBPB_diversity <- which(max(rowSums(abs(var_expt_levels[,3:6]))) ==
                            rowSums(abs(var_expt_levels[,3:6])) &
                              rowSums(abs(var_expt_levels[,1:2]))==0)

@@

```{r ss_novar, out.width="1\\linewidth", include=TRUE, fig.align="center", fig.cap=c("Steady states of the ecosystem with no trait diversity, i.e. identical trait values for the strains within a functional group."), echo=FALSE}
p1  <- plot_ss_result1(var_expt,
                result_index = no_diversity,
                filename_prefix = NULL,
                save_image_file = FALSE)
p1
```

@@

```{r ss_var, out.width="1\\linewidth", include=TRUE, fig.align="center", fig.cap=c("Steady states of the ecosystem with diversity in all functional groups"), echo=FALSE}
p2  <- plot_ss_result1(var_expt,
                result_index = max_diversty_all,
                filename_prefix = NULL,
                save_image_file = FALSE)
p2
```

@@

```{r ss_var2, out.width="1\\linewidth", include=TRUE, fig.align="center", fig.cap=c("Steady states of the ecosystem without any diversity (solid line) and with diversity in all three functional groups."), echo=FALSE}
p_overlay1 <- plot_ss_result2(var_expt[no_diversity,]$ss_res[[1]],
                             var_expt[max_diversty_all,]$ss_res[[1]],
                             xlims = c(-7, -1))
p_overlay1
```

@@

```{r CB_var1, out.width="1\\linewidth", include=TRUE, fig.align="center", fig.cap=c("Steady states of the ecosystem with diversity in only the cyanobacteria functional group."), echo=FALSE}
p3  <- plot_ss_result1(var_expt,
                result_index = max_only_CB_diversity,
                filename_prefix = NULL,
                save_image_file = FALSE)
p3
```

```{r CB_var2, out.width="1\\linewidth", include=TRUE, fig.align="center", fig.cap=c("Steady states of the ecosystem without any diversity (solid line) and with diversity in only the cyanobacteria functional group."), echo=FALSE}
p_overlay2 <- plot_ss_result2(var_expt[no_diversity,]$ss_res[[1]],
                             var_expt[max_only_CB_diversity,]$ss_res[[1]],
                             xlims = c(-7, -1))
p_overlay2
```

@@

```{r SBPB_var1, out.width="1\\linewidth", include=TRUE, fig.align="center", fig.cap=c("Steady states of the ecosystem with diversity in only the two sulphur bacteria functional groups."), echo=FALSE}
p4  <- plot_ss_result1(var_expt,
                result_index = max_only_SBPB_diversity,
                filename_prefix = NULL,
                save_image_file = FALSE)
p4

@@

```{r SBPB_var2, out.width="1\\linewidth", include=TRUE, fig.align="center", fig.cap=c("Steady states of the ecosystem without any diversity (solid line) and with diversity in only the sulphur bacteria functional group."), echo=FALSE}
p_overlay3 <- plot_ss_result2(var_expt[no_diversity,]$ss_res[[1]],
                             var_expt[max_only_SBPB_diversity,]$ss_res[[1]],
                             xlims = c(-7, -1))
p_overlay3
```
